---
title: "GAM_ESM_v2"
author: "Clara"
date: "`r Sys.Date()`"
output: html_document
---

## Setup
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# set root directory
setwd("C:/Users/cleme/Documents/Education/RUG/Thesis/EMA-mindfulness/Data")
```

## Clean everything up
```{r}
rm(list = ls())
```

## Import Libraries
```{r}
library(mgcv)
library(itsadug)
library(plyr)
library(ggplot2)
library(plotfunctions)
library(colorspace)
library (dplyr)
```

## Import data
```{r}
ESM_data <- read.csv("ESM_gams_g2_diff.csv")

```

# bs="re" -> include random intercepts and slopes
# bs="fs",m=1 -> add non-linear random effects
# BY parameter -> obtain separate non-linearities

```{r}
# need to make some columns of type factor to be able to include them as predictors

ESM_data$id2 <- as.factor(ESM_data$id2)
ESM_data$intervention <- as.factor(ESM_data$intervention)
ESM_data$subject <- as.factor(ESM_data$subject)
ESM_data$block <- as.factor(ESM_data$block)

```

## Create complex model
```{r}
m1 <- bam(db_stickiness_diff ~ s(assessmentDayPhase, by=intervention, k=6)
           + intervention
           + s(assessmentDayPhase, subject, by=block, bs="fs", m=1, k=6)
           , data=ESM_data, method="ML") 
summary(m1)

plot_smooth(m1, view="assessmentDayPhase", rug=F, plot_all="intervention")

# check autocorrelation
m1acf <- acf_resid(m1)

#check model fit
gam.check(m1)

#plot the model based predicted differences between interventions (summed effects)
plot_parametric(m1, pred=list(intervention=c("mindfulness", "fantasizing")))


# plot differences
plot_diff(m1, view="assessmentDayPhase", comp=list("intervention"=c("mindfulness", "fantasizing")), main="mindfulness vs fantasizing")

```


## Visualize raw data and check if it is in line with the model predictions
```{r}
avg2 <- ddply(ESM_data, c("group", "intervention", "assessmentDayPhase", "db_stickiness_diff"), summarise,
               N    = length(db_stickiness_diff),
               sd   = sd(db_stickiness_diff),
               se   = sd / sqrt(N))
plot2 <- ggplot(avg2, aes(y=db_stickiness_diff, x=assessmentDayPhase, color=intervention)) + geom_point(position = position_jitter(w=0.1,h=0))+ facet_grid(.~group) + ylab("Stickiness Difference peri-pre")+ xlab("Day") + labs(color = "Intervention") + geom_hline(yintercept=0) + scale_x_discrete(limits=c("1","2","3", "4", "5", "6", "7")) + geom_smooth() 

plot2

#ggsave(filename="Analysis/my_scripts/GAMs_new/visualisation/db_stickiness_rawDiff.png", plot=plot2, width=10, height=7)
```

## Create less complex model and compare
```{r}
# take out interaction
m2 <- bam(db_stickiness_diff ~ s(assessmentDayPhase, k=6)
           + intervention
           + s(assessmentDayPhase, subject, by=block, bs="fs", m=1, k=6)
           , data=ESM_data, method="ML") 
summary(m2)

# check autocorrelation
m2acf <- acf_resid(m2)

plot_smooth(m2, view="assessmentDayPhase", rug=F, plot_all="intervention")

compareML(m1, m2) #m2 better (AIC diff 0.45)


# take out main effect
m2b <- bam(db_stickiness_diff ~ s(assessmentDayPhase, k=6)
           # + intervention
           + s(assessmentDayPhase, subject, by=block, bs="fs", m=1, k=6)
           , data=ESM_data, method="ML") 
summary(m2b)

plot_smooth(m2b, view="assessmentDayPhase", rug=F, plot_all="intervention")

compareML(m2, m2b) #m2b slightly better (AIC diff 0.15)
```

## create intercept model and compare
```{r}
# intercept model
m.int <- bam(db_stickiness_diff ~ 1 + s(assessmentDayPhase, subject, by=block, bs="fs", m=1, k=6), data=ESM_data, method="ML")
summary(m.int)

compareML(m.int, m1) # intercept model vs complex model (AIC diff 4.89)
compareML(m.int, m2) # intercept model vs main effects model (AIC diff 5.34)
compareML(m1, m2) # m2 slightly better (AIC diff 0.45)
compareML(m2, m2b) #m2b slightly better (AIC diff 0.15)

# akaike 2011: AIC diff 2
```

## Investigate best model
```{r}
summary(m2b)
# Approximate significance of smooth terms:
#                                        edf Ref.df     F p-value    
# s(assessmentDayPhase)                 1.00      1 5.799  0.0162 *  

plot_smooth(m2b, view="assessmentDayPhase", rug=F, plot_all="intervention")

#check model fit
gam.check(m2b)

#plot the model based predicted differences between interventions (summed effects)
#plot_parametric(m2b, pred=list(intervention=c("mindfulness", "fantasizing")))

#plot differences
#plot_diff(m2b, view="assessmentDayPhase", comp=list("intervention"=c("mindfulness", "fantasizing")), main="mindfulness vs fantasizing")

```

